---
title: "Preprocessing for Thesis"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Loading Data In}
#setting up my dataset#
library(haven)
library(psych)
library(tibble)

df <- read_sav("pu_ssocs18.sav") #2017-2018 School Survey on Crime and Safety Dataset; find the hyperlink above#

#update#
basic_model <- df[, c("INCID18", "C0532", "C0534", "C0536")]

colnames(basic_model) <- c("total_incidents", "percentlowgrades", "percentgocollege", "percentacadimpt")

demographic_control <- df[, c("C0568", "STRATA", "C0560", "C0562")]

colnames(demographic_control) <- c("school_size", "demographics", "crime_live", "crime_area")

security <- df[, c("C0110", "C0112", "C0114", "C0116", "C0120", "C0121", "C0122", "C0125", "C0129", "C0134", "C0136", "C0138", "C0140", "C0139", "C0141", "C0143", "C0142", "C0144", "C0146", "C0150", "C0153")]

colnames(security) <- c("visitor_badge", "control_building", "control_ground", "require_check", "random_check", "lock_inside", "close_lunch", "random_sweep", "drug_testing", "wear_uniform", "dress_code", "school_lockers", "book_bags", "panic_button", "emergency_notif", "anonymous_report", "student_id", "staff_id", "security_camera", "radio", "prohibit_phone")

parental <- df[, c("")]

community <- 
  
educational_approaches <- 


                   
"crimeschool", "respondentyears", "classroomchanges", "randomsweeps", "prohibit_phone", "randommetal", "strictdresscode", "panicbutton") #adding variable names#
```

```{r Socio-Economic Variables}
#setting up my dataset#

```

```{r Security Variables}
#setting up my dataset#

```

```{r Community Variables}
#setting up my dataset#

```

```{r Parental Variables}
#setting up my dataset#

```

```{r Control Variables}
#setting up my dataset#

```

```{r Control Variables}
#setting up my dataset#

```






sub <- na.omit(sub) #just to be sure there are no NA values left#

checkzero_total_incidents <- sub$total_incidents == 0
table(checkzero_total_incidents) #there are 274 schools with zero incidents, out of all 2762 schools#

checkzero_percentlowgrades <- sub$percentlowgrades == 0
table(checkzero_percentlowgrades)#there are 182 schools with 0% low grades, out of all 2762 schools#

sub$percentnogocollege = 100 - sub$percentgocollege #reverse-recoding percentgocollege as percent no go college#

describe(sub$percentgocollege) #checking recoding#
describe(sub$percentnogocollege) #checking recoding#

sub$r.crimeschool = 4 - sub$crimeschool #reverse-coding crimeschool variable#
#now, higher level of r.crimeschool means higher level of crime#

table(sub$crimeschool) #checking recoding#
table(sub$r.crimeschool) #checking recoding#


sub$r.randomsweeps = 3 - sub$randomsweeps #reverse-coding randomsweeps variable#

sub$r.randomsweeps = ifelse(sub$r.randomsweeps == 2, 1, 0)

table(sub$randomsweeps) #checking recoding#
table(sub$r.randomsweeps) #checking recoding#

sub$r.prohibit_phone = 3 - sub$prohibit_phone #reverse-coding prohibit_phone variable#

sub$r.prohibit_phone = ifelse(sub$r.prohibit_phone == 2, 1, 0)

table(sub$prohibit_phone) #checking recoding#
table(sub$r.prohibit_phone) #checking recoding#

sub$r.randommetal = 3 - sub$randommetal #reverse-coding random_metal variable#

sub$r.randommetal = ifelse(sub$r.randommetal == 2, 1, 0)

table(sub$randommetal) #checking recoding#
table(sub$r.randommetal) #checking recoding#

sub$r.strictdresscode = 3 - sub$strictdresscode #reverse-coding strictdresscode variable#

sub$r.strictdresscode = ifelse(sub$r.strictdresscode == 2, 1, 0)

table(sub$strictdresscode) #checking recoding#
table(sub$r.strictdresscode) #checking recoding#

sub$r.panicbutton = 3 - sub$panicbutton #reverse-coding strictdresscode variable#
table(sub$panicbutton) #checking recoding#
table(sub$r.panicbutton) #checking recoding#

describe(sub$respondentyears)
sub$respondentyears[sub$respondentyears < 0] <- NA
describe(sub$respondentyears)
sd(sub$respondentyears, na.rm=TRUE)

describe(log1p(sub$respondentyears))
sd(log1p(sub$respondentyears), na.rm=TRUE)

describe(sub$classroomchanges)
sd(sub$classroomchanges)
```
 
```{r Descriptive Statistics}

descriptives <- sub[, c("total_incidents", "percentlowgrades", "percentgocollege")]

tocount <- sub[, c("r.randomsweeps", "r.prohibit_phone", "r.randommetal", "r.strictdresscode")]

tocount <- data.frame(tocount)

apply(tocount[-1], 2, table)

descriptives <- data.frame(descriptives)

library(stargazer)

stargazer(descriptives, covariate.labels = c("Total Incidents", "Percentage of Students with Low Grades", "Percentage of Students Unlikely to Attend College", "Random Sweeps in School", "Prohibition of Non-Academic Phone Use", "Random Metal Detector Checks on Students", "Strict Dress Code for Students"))

sub$t.total_incidents <- log1p(sub$total_incidents)

sub$t.percentlowgrades <- log1p(sub$percentlowgrades)

transform <- sub[, c("t.total_incidents", "t.percentlowgrades")]

transform <- data.frame(transform)

stargazer(transform, covariate.labels = c("Total Incidents", "Percentage of Students with Low Grades"))
```

```{r Save R Dataspace}
#setting up my dataset#

save.image(file = "my_work_space.RData")

```

